---
title: "Clean Data"
---

## Libraries

```{r}
library(tidyverse)
library(here)
library(brms)
library(tidybayes)
library(fastDummies)
library(ggokabeito)
library(patchwork)
library(cmdstanr)
```

## Functions

```{r}
source(here("size/scripts/size_helper_fcts.R"))
```

## Data

Combine the exported TagLab tables (1 tale per quadrat, 20 quadrats per site) and extract states.

### E2B

#### Read and set up data

```{r}
# read data
data_E2B <- combine_data(here("size/data/raw/2024_06_05_export/E2B"))

#combine multiple information per cell
data_E2B <- data_E2B %>%
  combine_multiple_entries()

#one df for each pair of years, re-set names of columns, get status info

data_16_18_E2B <- data_E2B %>%
  select(c(1:7)) %>%
  filter_at(vars(spec_1, spec_2), any_vars(!is.na(.))) %>%
  get_infos() %>%
  mutate(comparison = "2016_2018", site = "E2B")

data_18_20_E2B <- data_E2B %>%
  select(c(1, 5:10)) %>%
  filter_at(vars(spec_2, spec_3), any_vars(!is.na(.))) %>%
  rename(
    "spec_1" = "spec_2",
    "area_1" = "area_2",
    "ID_1" = "ID_2",
    "spec_2" = "spec_3",
    "area_2" = "area_3",
    "ID_2" = "ID_3"
  ) %>%
  get_infos() %>%
  mutate(comparison = "2018_2020", site = "E2B")


data_20_22_E2B <- data_E2B %>%
  select(c(1, 8:13)) %>%
  filter_at(vars(spec_3, spec_4), any_vars(!is.na(.))) %>%
  rename(
    "spec_1" = "spec_3",
    "area_1" = "area_3",
    "ID_1" = "ID_3",
    "spec_2" = "spec_4",
    "area_2" = "area_4",
    "ID_2" = "ID_4"
  ) %>%
  get_infos() %>%
  mutate(comparison = "2020_2022", site = "E2B")
```

#### Check data

Get overview of occurrences of each category

```{r}
data_16_18_E2B %>%
  group_by(spec_1, state) %>%
  count() %>%
  pivot_wider(names_from = state, values_from = n) %>%
  mutate(n_tot = rowSums(across(where(is.numeric))))

data_18_20_E2B %>%
  group_by(spec_1, state) %>%
  count() %>%
  pivot_wider(names_from = state, values_from = n) %>%
  mutate(n_tot = rowSums(across(where(is.numeric))))

data_20_22_E2B %>%
  group_by(spec_1, state) %>%
  count() %>%
  pivot_wider(names_from = state, values_from = n) %>%
  mutate(n_tot = rowSums(across(where(is.numeric))))
```

Plot relative change. Mainly to find unrealistic values

```{r}
data_16_18_E2B %>%
  ggplot(aes(x = rel_change)) +
  geom_histogram(bins = 100) +
  geom_vline(xintercept = 0, col = "darkred") +
  facet_wrap(~spec_1) +
  theme_minimal()

data_16_18_E2B %>%
  filter(rel_change > 20) # ok

data_18_20_E2B %>%
  ggplot(aes(x = rel_change)) +
  geom_histogram(bins = 100) +
  geom_vline(xintercept = 0, col = "darkred") +
  facet_wrap(~spec_1) +
  theme_minimal()

data_20_22_E2B %>%
  ggplot(aes(x = rel_change)) +
  geom_histogram(bins = 100) +
  geom_vline(xintercept = 0, col = "darkred") +
  facet_wrap(~spec_1) +
  theme_minimal()
```

Plot the categories

```{r}
data_16_18_E2B %>%
  ggplot(aes(x = area_1, col = state)) +
  geom_histogram(bins = 100, fill = NA) +
  facet_wrap(~spec_1) +
  scale_color_okabe_ito() +
  theme_minimal()
```

```{r}
data_18_20_E2B %>%
  ggplot(aes(x = area_1, col = state)) +
  geom_histogram(bins = 100, fill = NA) +
  facet_wrap(~spec_1) +
  scale_color_okabe_ito() +
  theme_minimal()
```

```{r}
data_20_22_E2B %>%
  ggplot(aes(x = area_1, col = state)) +
  geom_histogram(bins = 100, fill = NA) +
  facet_wrap(~spec_1) +
  scale_color_okabe_ito() +
  theme_minimal()
```

Now, three data sets are generated for each pair of years

1.  States
2.  Size per year

#### States

```{r}
dat_state_E2B <- bind_rows(data_16_18_E2B, data_18_20_E2B, data_20_22_E2B) %>%
  mutate(site = "E2B")
```

#### Size per year

```{r}
data_E2B <- data_E2B %>%
  select(spec_1, spec_2, spec_3, spec_4, area_1, area_2, area_3, area_4) %>%
  pivot_longer(area_1:area_4, names_to = "year", values_to = "area") %>%
  mutate(
    year = recode(
      year,
      "area_1" = "2016",
      "area_2" = "2018",
      "area_3" = "2020",
      "area_4" = "2022"
    ),
    site = "E2B",
    spec = ifelse(is.na(spec_1), spec_2, spec_1),
    spec = ifelse(is.na(spec), spec_3, spec),
    spec = ifelse(is.na(spec), spec_4, spec)
  ) %>%
  select(-spec_1, -spec_2, -spec_3, -spec_4) %>%
  filter(area > 0)
```

### Haapiti

#### Read and set up data

```{r}
# read data
data_HAA <- combine_data(here("size/data/raw/2024_06_05_export/HAA"))

#combine multiple information per cell
data_HAA <- data_HAA %>%
  combine_multiple_entries()

#one df for each pair of years, re-set names of columns, get status info

data_16_18_HAA <- data_HAA %>%
  select(c(1:7)) %>%
  filter_at(vars(spec_1, spec_2), any_vars(!is.na(.))) %>%
  get_infos() %>%
  mutate(comparison = "2016_2018", site = "HAA")

data_18_20_HAA <- data_HAA %>%
  select(c(1, 5:10)) %>%
  filter_at(vars(spec_2, spec_3), any_vars(!is.na(.))) %>%
  rename(
    "spec_1" = "spec_2",
    "area_1" = "area_2",
    "ID_1" = "ID_2",
    "spec_2" = "spec_3",
    "area_2" = "area_3",
    "ID_2" = "ID_3"
  ) %>%
  get_infos() %>%
  mutate(comparison = "2018_2020", site = "HAA")


data_20_22_HAA <- data_HAA %>%
  select(c(1, 8:13)) %>%
  filter_at(vars(spec_3, spec_4), any_vars(!is.na(.))) %>%
  rename(
    "spec_1" = "spec_3",
    "area_1" = "area_3",
    "ID_1" = "ID_3",
    "spec_2" = "spec_4",
    "area_2" = "area_4",
    "ID_2" = "ID_4"
  ) %>%
  get_infos() %>%
  mutate(comparison = "2020_2022", site = "HAA")
```

#### Check data

Get overview of occurrences of each category

```{r}
data_16_18_HAA %>%
  group_by(spec_1, state) %>%
  count() %>%
  pivot_wider(names_from = state, values_from = n) %>%
  mutate(n_tot = rowSums(across(where(is.numeric))))

data_18_20_HAA %>%
  group_by(spec_1, state) %>%
  count() %>%
  pivot_wider(names_from = state, values_from = n) %>%
  mutate(n_tot = rowSums(across(where(is.numeric))))

data_20_22_HAA %>%
  group_by(spec_1, state) %>%
  count() %>%
  pivot_wider(names_from = state, values_from = n) %>%
  mutate(n_tot = rowSums(across(where(is.numeric))))
```

Plot relative change. Mainly to find unrealistic values

```{r}
data_16_18_HAA %>%
  ggplot(aes(x = rel_change)) +
  geom_histogram(bins = 100) +
  geom_vline(xintercept = 0, col = "darkred") +
  facet_wrap(~spec_1) +
  theme_minimal()

data_16_18_HAA %>%
  filter(rel_change > 20) # ok

data_18_20_HAA %>%
  ggplot(aes(x = rel_change)) +
  geom_histogram(bins = 100) +
  geom_vline(xintercept = 0, col = "darkred") +
  facet_wrap(~spec_1) +
  theme_minimal()

data_20_22_HAA %>%
  ggplot(aes(x = rel_change)) +
  geom_histogram(bins = 100) +
  geom_vline(xintercept = 0, col = "darkred") +
  facet_wrap(~spec_1) +
  theme_minimal()
```

Plot the categories

```{r}
data_16_18_HAA %>%
  ggplot(aes(x = area_1, col = state)) +
  geom_histogram(bins = 100, fill = NA) +
  facet_wrap(~spec_1) +
  scale_color_okabe_ito() +
  theme_minimal()
```

```{r}
data_18_20_HAA %>%
  ggplot(aes(x = area_1, col = state)) +
  geom_histogram(bins = 100, fill = NA) +
  facet_wrap(~spec_1) +
  scale_color_okabe_ito() +
  theme_minimal()
```

```{r}
data_20_22_HAA %>%
  ggplot(aes(x = area_1, col = state)) +
  geom_histogram(bins = 100, fill = NA) +
  facet_wrap(~spec_1) +
  scale_color_okabe_ito() +
  theme_minimal()
```

Now, three data sets are generated for each pair of years

1.  States
2.  Size per year

#### States

```{r}
dat_state_HAA <- bind_rows(data_16_18_HAA, data_18_20_HAA, data_20_22_HAA) %>%
  mutate(site = "HAA")
```

#### Size per year

```{r}
data_HAA <- data_HAA %>%
  select(spec_1, spec_2, spec_3, spec_4, area_1, area_2, area_3, area_4) %>%
  pivot_longer(area_1:area_4, names_to = "year", values_to = "area") %>%
  mutate(
    year = recode(
      year,
      "area_1" = "2016",
      "area_2" = "2018",
      "area_3" = "2020",
      "area_4" = "2022"
    ),
    site = "HAA",
    spec = ifelse(is.na(spec_1), spec_2, spec_1),
    spec = ifelse(is.na(spec), spec_3, spec),
    spec = ifelse(is.na(spec), spec_4, spec)
  ) %>%
  select(-spec_1, -spec_2, -spec_3, -spec_4) %>%
  filter(area > 0)
```

### Combine & Save

```{r}
dat_state <- bind_rows(dat_state_E2B, dat_state_HAA) %>%
  filter(spec_1 == "Pocillopora")

dat_size <- bind_rows(data_E2B, data_HAA) %>%
  filter(spec == "Pocillopora")

write.csv(dat_state, file = here("size/data/dat_state.csv"), row.names = F)
write.csv(dat_size, file = here("size/data/dat_size.csv"), row.names = F)
```