---
title: "Checks n numbers"
format: html
---

```{r}
library(tidyverse)
```

## Get data

```{r}
dat <- readRDS(here("long_term/data/clean/clean_MooreaMPA.rds"))
dat_meta <- read.csv(here("long_term/data/clean/moorea_mpa_site_metadata.csv"))
```

- Filter to only coral, outer slope, after 2018

```{r}
dat_c <- dat %>%
  filter(benthic_group == "coral", habitat == "Outer slope", year > 2018)
```


Select 3 main genera

```{r}
top_3_corals <- c("Pocillopora", "Montipora", "Acropora")
```

bin other genera into "others", re-calculate cover

```{r}
dat_c <- dat_c %>%
  left_join(dat_meta, by = "marine_area") %>%
  mutate(genus = ifelse(substrate %in% top_3_corals, substrate, "Other")) %>%
  dplyr::select(-substrate) %>%
  group_by(across(c(-proportion))) %>%
  summarise(proportion = sum(proportion, na.rm = T)) %>%
  ungroup() %>%
  pivot_wider(names_from = genus, values_from = proportion, values_fill = 0) %>%
  dplyr::select(-benthic_group, -habitat) %>%
  pivot_longer(
    cols = 10:ncol(.),
    names_to = "genus",
    values_to = "proportion"
  ) %>%
  mutate(genus = fct_relevel(genus, "Other", after = Inf)) #set others to end
```


## Absolute/observed Cover: Overviews for in-text numbers & examples

Over all sites, for 2019 & 2020

```{r}
dat_c %>%
  filter(year %in% 2019:2020, genus != "Other") %>%
  group_by(genus, year) %>%
  summarise(
    mean_cover = mean(proportion * 100),
    sd = sd(proportion * 100),
    n = n(),
    se = round(sd / sqrt(n), 2)
  ) %>%
  mutate(mean_cover = round(mean_cover, 2), se = round(se, 2)) %>%
  dplyr::select(genus, year, mean_cover, se)
```


### Points and percent

Extract points per year

```{r}
#Acropora
dat_c %>%
  mutate(points_coral = proportion * 50, points_total = 50) %>%
  filter(genus == "Acropora") %>%
  group_by(year) %>%
  summarise(sum_points = sum(points_coral), mean_points = mean(points_coral))

#Montipora
dat_c %>%
  mutate(points_coral = proportion * 50, points_total = 50) %>%
  filter(genus == "Montipora") %>%
  group_by(year) %>%
  summarise(sum_points = sum(points_coral), mean_points = mean(points_coral))
```


At Taotaha

```{r}
dat_c %>%
  mutate(points_coral = proportion * 50, points_total = 50) %>%
  group_by(genus, year) %>%
  summarise(
    points_coral = sum(points_coral),
    points_total = sum(points_total),
    percent = round(100 * points_coral / points_total, 1)
  )
```

Taotaha Montipora

```{r}
dat_c %>%
  filter(marine_area == "Taotaha", genus == "Montipora") %>%
  mutate(points_coral = proportion * 50, points_total = 50) %>%
  group_by(year) %>%
  summarise(
    points_coral = sum(points_coral),
    points_total = sum(points_total),
    percent = round(100 * points_coral / points_total, 1)
  )
```



## Model Delta

Prepare data for model

Get DHW data at to data. Names of sites differ and needs to be aligned

```{r}
dat_genus_delta <- dat_c

dat_dhw <- read.csv(here("dhw/data/dat_sites_maxdhw_mur.csv"))


# rename dat_coral_delta sites

dat_genus_delta <- dat_genus_delta %>%
  mutate(
    marine_area = case_when(
      marine_area == "Haapiti" ~ "Ha'apiti",
      marine_area == "Maatea" ~ "Ma'atea",
      marine_area == "Pihaena" ~ "Piha'ena",
      marine_area == "Temae" ~ "Tema'e",
      .default = marine_area
    )
  )

dat_dhw <- dat_dhw %>%
  mutate(site = ifelse(site == "Entre 2 baies", "Entre 2 baies", site))

dat_genus_delta <- dat_genus_delta %>%
  dplyr::select(-coast) %>%
  left_join(dat_dhw, by = c("marine_area" = "site"))
```


- 25 m every 50 cm per transect =\> 50 points
- remove "Other", mix of genera, no common response expected

```{r}
dat_genus_delta <- dat_genus_delta %>%
  mutate(counts = as.integer(proportion * 50)) %>%
  filter(genus != "Other") %>%
  droplevels()

#ensure year is factor
dat_genus_delta <- dat_genus_delta %>%
  mutate(
    f_year = paste0("f", year) %>%
      factor()
  )
```


Save as rds

```{r}
write_rds(dat_genus_delta, here("long_term/data/clean/dat_genus_delta.rds"))
```


## Plot Cover top 5 corals 2005 to 2023

Get data

```{r}
#take original data
dat_allgenera <- dat %>%
  left_join(dat_meta, by = "marine_area") %>%
  filter(benthic_group %in% c("coral", "macro_alga"), habitat == "Outer slope")

top_5_corals <- dat_allgenera %>%
  filter(benthic_group == "coral") %>%
  group_by(year, marine_area, substrate) %>%
  summarise(proportion = mean(proportion)) %>% #mean coral cover per site
  group_by(substrate) %>%
  summarise(
    proportion = mean(proportion) %>%
      round(2)
  ) %>%
  arrange(desc(proportion)) %>%
  slice(1:5) %>%
  pull(substrate)


dat_allgenera <- dat_allgenera %>%
  mutate(
    genus = ifelse(benthic_group == "macro_alga", "Macroalgae", substrate)
  ) %>%
  mutate(
    genus = ifelse(
      genus %in% c(top_5_corals, "Macroalgae"),
      genus,
      "Other Corals"
    )
  ) %>%
  dplyr::select(-substrate) %>%
  group_by(across(c(-proportion))) %>%
  summarise(proportion = sum(proportion, na.rm = T)) %>%
  ungroup() %>%
  pivot_wider(names_from = genus, values_from = proportion, values_fill = 0) %>%
  dplyr::select(-benthic_group, -habitat) %>%
  pivot_longer(
    cols = 10:ncol(.),
    names_to = "genus",
    values_to = "proportion"
  ) %>%
  mutate(genus = fct_relevel(genus, "Other Corals", after = Inf)) %>% #set others to  last
  mutate(genus = fct_relevel(genus, "Macroalgae", after = 0L)) #set Macroalgae to first

# adapt site names
dat_allgenera <- dat_allgenera %>%
  mutate(
    marine_area = case_when(
      marine_area == "Entre 2 baies" ~ "Entre deux baies",
      marine_area == "Haapiti" ~ "Ha'apiti",
      marine_area == "Maatea" ~ "Ma'atea",
      marine_area == "Pihaena" ~ "Piha'ena",
      marine_area == "Temae" ~ "Tema'e",
      .default = marine_area
    )
  )
```

```{r}
cols_allgenera <- c(
  cols_genera[1:3],
  "Other Corals" = "grey70",
  "Porites" = brewer.pal(11, "Set3")[7],
  "Astrea" = brewer.pal(11, "Set3")[8],
  "Macroalgae" = "darkgreen"
)


plot_coral_cover <- dat_allgenera %>%
  group_by(marine_area, year, coast, genus) %>%
  summarise(cover_perc = mean(proportion * 100, na.rm = T)) %>%
  ungroup() %>%
  mutate(coast = factor(coast, levels = c("West", "North", "East"))) %>%
  ggplot() +

  geom_area(aes(fill = genus, x = year, y = cover_perc)) +

  labs(x = "Year", y = "Benthic cover (%)") +
  scale_fill_manual(values = cols_allgenera) +
  guides(fill = guide_legend(title = "Genus")) +

  facet_ragged_cols(vars(marine_area), vars(coast)) +
  theme_minimal() +
  theme(legend.position = "bottom") +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.ticks = element_line(colour = "black", size = 0.2),
    panel.border = element_rect(fill = NA)
  )
plot_coral_cover
```

```{r}
ggsave(
  filename = "SI_coralgenera_cover.pdf",
  plot = plot_coral_cover,
  path = here("long_term/plots/"),
  width = 20,
  height = 15,
  units = "cm"
)
```
