---
title: "Model: Bleaching Mortality and Recovery"
---

## Libraries & Functions

```{r}
library(tidyverse)
library(here)
library(RColorBrewer)
library(brms)
library(tidybayes)
library(patchwork)
library(bayesplot)
library(ggragged)
library(ggridges)
library(janitor)
library(htmlTable)
library(viridis)
library(ggpp)
library(ggrepel)
library(performance)

# get functions

source(here("long_term/scripts/model_helper_fcts.R"))
```


### Model recovery

-  binomial distribution to model counts with resp. coral genera alogn 50 points per transect for all years incl. 2019

```{r}
m_ccover_genus <- brm(
  counts |
    trials(50) ~
    max_dhw *
    f_year *
    genus +
    (1 | marine_area) +
    (1 | genus) +
    (1 | f_year:marine_area:genus),
  data = dat_genus_delta,
  family = binomial(),
  iter = 4000,
  prior = c(
    prior(normal(0, 10), class = b),
    prior(normal(0, 10), class = Intercept)
  ),
  warmup = 2000,
  chains = 4,
  cores = 4,
  threads = threading(2, static = T),
  control = list(adapt_delta = 0.99, max_treedepth = 13),
  seed = 123,
  file = here("long_term/models/m_cover_genus"),
  backend = "cmdstanr"
)
```

Model checks

```{r}
summary(m_ccover_genus) #Rhats ok, 5 div trans
pp_check(m_ccover_genus, ndraws = 100) # good

r2_bayes(m_ccover_genus)
# Conditional R2: 0.890 (95% CI [0.876, 0.901])
# Marginal R2: 0.573 (95% CI [0.071, 0.778])
```

#### Impact AHS_max

get values

```{r}
min(dat_genus_delta$max_dhw) %>% round(2) # 5.76
mean(dat_genus_delta$max_dhw) %>% round(2) # 6.06
max(dat_genus_delta$max_dhw) %>% round(2) # 6.26
```

Predict expected values for impact of AHS_max (along sequence from lowest to highest value) w/o site-specific impact

```{r}
# maake new data frame

new_data_ahs_max <- dat_genus_delta %>%
  dplyr::select(genus, f_year) %>%
  distinct() %>%
  expand_grid(
    max_dhw = c(
      min(dat_genus_delta$max_dhw),
      mean(dat_genus_delta$max_dhw),
      max(dat_genus_delta$max_dhw)
    )
  )

# and use to predict

pred_ahs_max <- epred_draws(
  m_ccover_genus,
  newdata = new_data_ahs_max,
  re_formula = NA
) %>%
  ungroup() %>%
  dplyr::select(.draw, max_dhw, genus, f_year, .epred) %>%
  pivot_wider(names_from = f_year, values_from = .epred) %>%
  clean_names() %>%
  mutate(
    perc_2019 = 100 * (f2019 / 50),
    perc_2020 = 100 * (f2020 / 50),
    perc_2021 = 100 * (f2021 / 50),
    perc_2022 = 100 * (f2022 / 50),
    perc_2023 = 100 * (f2023 / 50)
  ) %>%
  mutate(
    delta_2020 = 100 * perc_2020 / perc_2019,
    delta_2021 = 100 * perc_2021 / perc_2019,
    delta_2022 = 100 * perc_2022 / perc_2019,
    delta_2023 = 100 * perc_2023 / perc_2019
  ) %>%
  ungroup() %>%
  dplyr::select(
    max_dhw,
    genus,
    delta_2020,
    delta_2021,
    delta_2022,
    delta_2023,
    draw
  ) %>%
  pivot_longer(
    delta_2020:delta_2023,
    names_to = "year",
    values_to = "percent"
  ) %>%
  mutate(
    year = substr(year, 7, 10) %>%
      as.numeric()
  )


#summary with median and CIs
pred_ahs_maxS <- pred_ahs_max %>%
  group_by(year, max_dhw, genus) %>%
  summarise(
    median = median(percent),
    l_89 = quantile(percent, probs = (1 - 0.89) / 2),
    u_89 = quantile(percent, probs = 1 - (1 - 0.89) / 2),
    l_95 = quantile(percent, probs = (1 - 0.95) / 2),
    u_95 = quantile(percent, probs = 1 - (1 - 0.95) / 2),
  ) %>%

  ungroup() %>%
  add_row(expand.grid(
    year = 2019,
    max_dhw = pred_ahs_max$max_dhw %>% unique(),
    genus = pred_ahs_max$genus %>% unique(),
    median = 100
  )) %>%
  mutate(
    AHS = case_when(
      max_dhw < 6 ~ "min",
      max_dhw > 6.2 ~ "max",
      .default = "mean"
    )
  ) %>%
  mutate(
    genus = fct_rev(genus),
    AHS = factor(AHS, levels = c("min", "mean", "max"))
  )

#random draws of distribution
set.seed(123)
sample_draws <- sample(1:8000, size = 50)

pred_ahs_max_100 <- pred_ahs_max %>%
  group_by(max_dhw, genus, year) %>%
  filter(draw %in% sample_draws) %>%
  ungroup() %>%
  add_row(expand.grid(
    year = 2019,
    max_dhw = pred_ahs_max$max_dhw %>% unique(),
    genus = pred_ahs_max$genus %>% unique(),
    draw = sample_draws,
    percent = 100
  )) %>%
  mutate(group = paste(max_dhw, draw, sep = "_")) %>%
  mutate(
    AHS = case_when(
      max_dhw < 6 ~ "min",
      max_dhw > 6.2 ~ "max",
      .default = "mean"
    )
  ) %>%
  mutate(
    genus = fct_rev(genus),
    AHS = factor(AHS, levels = c("min", "mean", "max"))
  )

col_ahs <- c("max" = "#FC4E07", "mean" = "#E7B800", "min" = "#00AFBB")


plot_change_ahs_max <- pred_ahs_max_100 %>%
  ggplot(aes(x = year)) +
  geom_hline(yintercept = 100, linetype = "11", col = "grey") +
  geom_line(
    aes(y = percent, col = AHS, group = group),
    linewidth = .2,
    alpha = .5
  ) +
  geom_line(
    data = pred_ahs_maxS,
    aes(y = median, group = AHS),
    linewidth = 2,
    col = "white"
  ) +
  geom_line(
    data = pred_ahs_maxS,
    aes(y = median, col = AHS, group = AHS),
    linewidth = 1
  ) +
  geom_point(
    data = pred_ahs_maxS,
    aes(y = median, fill = AHS, group = AHS),
    shape = 21,
    stroke = .7,
    col = "white",
    size = 2
  ) +
  scale_colour_manual(name = expression(AHS[max]), values = col_ahs) +
  scale_fill_manual(name = expression(AHS[max]), values = col_ahs) +
  facet_grid(~genus) +
  labs(y = expression("Relative cover (%)"), x = NULL) +
  theme_andi() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    plot.title = element_text(hjust = 0.5, face = "bold"),
    strip.text = element_text(face = "italic"),
    legend.position = "bottom"
  )
plot_change_ahs_max


ggsave(
  filename = "plot_change_ahs_max.pdf",
  plot = plot_change_ahs_max,
  path = here("long_term/plots/"),
  height = 15 * 0.8,
  width = 25 * 0.8,
  units = "cm"
)
```

#### Summary table

```{r}
pred_ahs_maxS %>%
  filter(year != 2019) %>%
  mutate(
    m_SI = paste0(
      round(median, 0),
      " [",
      round(l_95, 0),
      ", ",
      round(u_95, 0),
      "]"
    )
  ) %>%
  select(year, AHS, genus, m_SI) %>%
  pivot_wider(names_from = genus, values_from = m_SI) %>%
  select(!year) %>%
  htmlTable(rgroup = 2020:2023, n.rgroup = rep(3, 4), rnames = FALSE)
```

#### Impact marine_area

```{r}
new_data_marine_area <- dat_genus_delta %>%
  dplyr::select(genus, f_year, marine_area) %>%
  distinct() %>%
  left_join(
    dat_genus_delta %>%
      select(marine_area, max_dhw) %>%
      distinct(),
    by = "marine_area"
  )


pred_marine_area <- epred_draws(
  m_ccover_genus,
  newdata = new_data_marine_area,
  re_formula = NULL
) %>% # include population random effect
  ungroup() %>%
  dplyr::select(.draw, marine_area, genus, f_year, .epred, max_dhw) %>%
  pivot_wider(names_from = f_year, values_from = .epred) %>%
  clean_names() %>%
  mutate(
    perc_2019 = 100 * (f2019 / 50),
    perc_2020 = 100 * (f2020 / 50),
    perc_2021 = 100 * (f2021 / 50),
    perc_2022 = 100 * (f2022 / 50),
    perc_2023 = 100 * (f2023 / 50)
  ) %>%
  mutate(
    delta_2020 = 100 * perc_2020 / perc_2019,
    delta_2021 = 100 * perc_2021 / perc_2019,
    delta_2022 = 100 * perc_2022 / perc_2019,
    delta_2023 = 100 * perc_2023 / perc_2019
  ) %>%
  ungroup() %>%
  dplyr::select(
    marine_area,
    genus,
    delta_2020,
    delta_2021,
    delta_2022,
    delta_2023,
    draw,
    max_dhw
  ) %>%
  pivot_longer(
    delta_2020:delta_2023,
    names_to = "year",
    values_to = "percent"
  ) %>%
  mutate(
    year = substr(year, 7, 10) %>%
      as.numeric()
  )


bin_ahs <- data.frame(
  max_dhw = pred_marine_area$max_dhw %>%
    unique() %>%
    sort(),
  bin = c(1, 1, 1, 1, 1, 2, 2, 2, 2, 3, 3, 3, 3)
) %>%
  group_by(bin) %>%
  mutate(
    min = round(min(max_dhw), 2),
    max = round(max(max_dhw), 2),
    gr_n = paste(min, max, sep = " - ")
  ) %>%
  ungroup() %>%
  mutate(
    cols = c(
      "#E69F00",
      "#56B4E9",
      "#009E73",
      "#999999",
      "#CC79A7",
      "#E69F00",
      "#56B4E9",
      "#009E73",
      "#999999",
      "#E69F00",
      "#56B4E9",
      "#009E73",
      "#999999"
    )
  ) %>%
  select(max_dhw, gr_n, cols)


pred_marine_areaS <- pred_marine_area %>%
  group_by(year, marine_area, genus, max_dhw) %>%
  summarise(
    median = median(percent),
    l_89 = quantile(percent, probs = (1 - 0.89) / 2),
    u_89 = quantile(percent, probs = 1 - (1 - 0.89) / 2),
    l_95 = quantile(percent, 0.025, na.rm = T),
    u_95 = quantile(percent, 0.975, na.rm = T),
  ) %>%
  mutate(
    gr_ahs = cut(
      max_dhw,
      #  breaks = quantile(max_dhw, probs = seq(0, 1, length.out = 3)),
      breaks = c(5.765000, 6.164286, 6.264714),
      include.lowest = TRUE
    )
  ) %>%
  left_join(bin_ahs, by = "max_dhw") %>%
  mutate(
    genus = fct_rev(genus),
    gr_n = paste0("AHS[max] * ': ' * ", gr_n, " * ~degree * C ~ weeks"),
    genus_orig = genus
  )

levels(pred_marine_areaS$genus) <- c(
  "Pocillopora" = expression(italic("Pocillopora")),
  "Montipora" = expression(italic("Montipora")),
  "Acropora" = expression(italic("Acropora"))
)


# get colours
cols_sites <- pred_marine_areaS %>%
  ungroup() %>%
  select(marine_area, cols) %>%
  distinct() %>%
  as.vector()

cols_sites <- setNames(as.character(cols_sites$cols), cols_sites$marine_area)
```

```{r}
datlab <- pred_marine_areaS %>%
  filter(year == 2020)


plot_change_marine_area <- pred_marine_areaS %>%
  ggplot(aes(x = year, group = marine_area)) +
  geom_hline(yintercept = 100, linetype = "11", colour = "grey") +
  geom_errorbar(
    aes(ymin = l_89, ymax = u_89, col = marine_area),
    width = 0,
    position = position_dodge(width = .3)
  ) +
  geom_line(
    aes(y = median, colour = marine_area),
    position = position_dodge(width = .3)
  ) +
  geom_point(
    aes(y = median, colour = marine_area),
    position = position_dodge(width = .3),
    size = .8
  ) +
  geom_text_repel(
    data = datlab,
    aes(
      color = marine_area,
      label = marine_area,
      y = median
    ),
    direction = "y",
    hjust = "right",
    size = 3,
    force = 10,
    segment.size = .5,
    segment.alpha = .5,
    segment.linetype = "11",
    box.padding = .4,
    segment.curvature = 0.4,
    segment.ncp = 3,
    segment.angle = 20,
    position = position_dodgenudge_to(width = .3, x = 2018.4),
  ) +
  scale_x_continuous(breaks = seq(2020, 2023, by = 1), limits = c(2018.2, NA)) +
  scale_colour_manual(values = cols_sites) +
  labs(y = expression("Relative cover (%)"), x = NULL) +
  facet_grid(genus ~ gr_n, labeller = label_parsed, scales = "free") +
  theme_andi() +
  theme(
    legend.position = "none",
    panel.background = element_rect(colour = "black", fill = 'transparent'),
    strip.text.y = element_text(face = "italic")
  )


plot_change_marine_area

ggsave(
  filename = "plot_change_marine_area.pdf",
  plot = plot_change_marine_area,
  path = here("long_term/plots/"),
  width = 28 * 0.8,
  height = 25 * 0.8,
  units = "cm"
)
```

#### Table over all sites

```{r}
pred_marine_area %>%
  filter(year != 2019) %>%
  group_by(genus, year) %>%
  summarise(
    cover = median(percent),
    l_95 = quantile(percent, 0.025, na.rm = T),
    u_95 = quantile(percent, 0.975, na.rm = T)
  ) %>%
  mutate(
    m_SI = paste0(
      round(cover, 0),
      " [",
      round(l_95, 0),
      ", ",
      round(u_95, 0),
      "]"
    )
  ) %>%
  ungroup() %>%
  select(genus, year, m_SI) %>%
  pivot_wider(names_from = year, values_from = m_SI) %>%
  htmlTable(
    rnames = FALSE
  )
```

#### Table for SI

```{r}
pred_marine_areaS %>%
  filter(year != 2019) %>%
  mutate(
    m_SI = paste0(
      round(median, 0),
      " [",
      round(l_95, 0),
      ", ",
      round(u_95, 0),
      "]"
    )
  ) %>%
  ungroup() %>%
  mutate(genus_year = paste(genus_orig, year, sep = "_")) %>%
  select(marine_area, genus_year, m_SI) %>%
  pivot_wider(names_from = genus_year, values_from = m_SI) %>%
  htmlTable(
    cgroup = c("", 2020:2023),
    n.cgroup = c(1, rep(3, 4)),
    rnames = FALSE
  )
```

## Model immediate mortality due to DHW in 2019

### Model

As above, this time only 2019 and 2020 data an dincluding year as interactive effect


```{r}
m_ccover_genus_19_dhw <- brm(
  counts | trials(50) ~
    max_dhw * genus * factor(year) + (1 + factor(year) | marine_area),
  data = dat_genus_delta %>%
    filter(year %in% 2019:2020),
  family = binomial(),
  iter = 4000,
  warmup = 2000,
  prior = c(
    prior(normal(0, 10), class = b),
    prior(normal(0, 10), class = Intercept)
  ),
  chains = 4,
  cores = 4,
  threads = threading(2, static = T),
  seed = 123,
  file = here("long_term/models/m_ccover_genus_19_dhw"),
  backend = "cmdstanr"
)
```

### Checks

```{r}
summary(m_ccover_genus_19_dhw) #no div trans
pp_check(m_ccover_genus_19_dhw, ndraws = 100) #good
bayes_R2(m_ccover_genus_19_dhw) #0.85

r2_bayes(m_ccover_genus_19_dhw)
# Conditional R2: 0.849 (95% CI [0.824, 0.871])
# Marginal R2: 0.742 (95% CI [0.669, 0.793])
```

### Line plot

Predict delta for each genus along sequence of max_DHW values

```{r}
new_data_dhw_gen <- expand_grid(
  max_dhw = seq(
    min(dat_genus_delta$max_dhw) - 0.1,
    max(dat_genus_delta$max_dhw) + 0.1,
    length = 100
  ),
  genus = unique(dat_genus_delta$genus),
  year = 2019:2020
)

pred_dhw_gen <- epred_draws(
  m_ccover_genus_19_dhw,
  newdata = new_data_dhw_gen,
  re_formula = NA
) %>%
  ungroup() %>%
  dplyr::select(max_dhw, .draw, .epred, year, genus) %>%
  pivot_wider(names_from = year, values_from = .epred) %>%
  mutate(delta = (`2020` - `2019`) / `2019`)


pred_dhw_genS <- pred_dhw_gen %>%
  group_by(max_dhw, genus) %>%
  summarise(
    median = median(delta, na.rm = T),
    l95 = quantile(delta, 0.025, na.rm = T),
    u95 = quantile(delta, 0.975, na.rm = T),

    l80 = quantile(delta, 0.1, na.rm = T),
    u80 = quantile(delta, 0.9, na.rm = T),

    l60 = quantile(delta, 0.2, na.rm = T),
    u60 = quantile(delta, 0.8, na.rm = T)
  ) %>%
  ungroup()

dat_genus_deltaS <- dat_genus_delta %>%
  dplyr::select(marine_area, proportion, year, coast, max_dhw, genus) %>%
  filter(year %in% c(2019:2020)) %>%
  pivot_wider(names_from = year, values_from = proportion, values_fn = mean) %>%
  mutate(delta = (`2020` - `2019`) / `2019`)


plot_dhw <- pred_dhw_genS %>%
  mutate(genus = fct_rev(genus)) %>%
  ggplot(aes(x = max_dhw)) +
  geom_hline(yintercept = 0, linetype = "11", col = "grey") +
  geom_ribbon(
    aes(ymin = l95 * 100, ymax = u95 * 100, fill = genus),
    alpha = .3
  ) +
  geom_ribbon(
    aes(ymin = l80 * 100, ymax = u80 * 100, fill = genus),
    alpha = .3
  ) +
  geom_ribbon(
    aes(ymin = l60 * 100, ymax = u60 * 100, fill = genus),
    alpha = .3
  ) +
  geom_line(
    aes(y = median * 100, col = genus),
    linewidth = 1,
    show.legend = F
  ) +
  geom_point(
    data = dat_genus_deltaS,
    aes(x = max_dhw, y = delta * 100, fill = genus),
    size = 2.5,
    shape = 21,
    col = "white",
    show.legend = F
  ) +
  scale_colour_manual(values = cols_genera, name = NULL) +
  scale_fill_manual(values = cols_genera, name = NULL) +
  facet_wrap(~genus, ncol = 1) +
  labs(
    y = expression(Delta ~ "Relative cover (%" ~ yr^-1 * ")"),
    x = expression(AHS[max]),
    tag = "A"
  ) +
  theme_andi() +
  theme(
    legend.position = "none",
    plot.title = element_text(face = "bold", hjust = 0.5),
    panel.background = element_rect(colour = "black", fill = 'transparent'),
    strip.text = element_text(face = "italic")
  )

plot_dhw
```

#### Mortality over all site

For info within text

```{r}
pred_dhw_gen %>%
  group_by(genus) %>%
  summarise(
    median = median(delta * 100, na.rm = T),
    l_95 = quantile(delta * 100, 0.025, na.rm = T),
    u_95 = quantile(delta * 100, 0.975, na.rm = T)
  ) %>%
  mutate_if(is.numeric, txtRound, 0) %>%
  htmlTable(rnames = FALSE)
```

### Slope

Assuming linear relationship over whole course of max_DHW calculate slope between lowest max_DHW value and highest. Calculate this slope for each genus

```{r}
slope_gen <- epred_draws(
  m_ccover_genus_19_dhw,
  re_formula = NA,
  newdata = expand_grid(
    max_dhw = c(min(dat_genus_delta$max_dhw), max(dat_genus_delta$max_dhw)),
    year = 2019:2020,
    genus = dat_genus_delta$genus %>% unique()
  )
) %>%
  ungroup() %>%
  dplyr::select(max_dhw, year, .epred, .draw, genus) %>%
  pivot_wider(names_from = year, values_from = .epred) %>%
  mutate(delta = (`2020` - `2019`) / `2019`) %>%
  dplyr::select(delta, max_dhw, .draw, genus) %>%
  pivot_wider(names_from = max_dhw, values_from = delta) %>%
  mutate(
    slope_val = (`6.26471428571429` - `5.765`) / (6.26471428571429 - 5.765)
  ) %>%
  mutate(
    genus = fct_reorder(.f = genus, .x = slope_val, .fun = mean, .desc = T)
  )
```

Calculate how proportion of distribution \> 0

```{r}
dat_above0 <- slope_gen %>%
  group_by(genus) %>%
  summarize(
    above_0 = above_0(slope_val),
    below_0 = 100 * round(1 - above_0, 4)
  ) %>%
  mutate(above_0 = paste0(round(above_0 * 100, 0), "%")) %>%
  mutate(above_0 = ifelse(above_0 == "0%", "<1%", above_0))
```

### Plot slopes

```{r}
plot_slope_gen <- slope_gen %>%
  ggplot(aes(y = genus, x = slope_val * 100)) +
  stat_halfeye(aes(fill = genus), .width = c(0.8, 0.95), ) +
  geom_vline(xintercept = 0, linetype = "11") +
  geom_text(
    data = dat_above0,
    aes(x = 8, label = above_0),
    hjust = 0,
    vjust = -2
  ) +
  labs(
    #   x = expression(Delta ~ "% Relative cover" ~ °C^-1 ~ "week"^-1),
    x = expression(Delta ~ "% Relative cover" ~ degree * C^-1 ~ week^-1),
    y = NULL,
    tag = "B"
  ) +
  scale_fill_manual(values = cols_genera, name = NULL) +
  coord_cartesian(xlim = c(-300, 200)) +
  theme_andi() +
  theme(legend.position = "none", axis.text.y = element_text(face = 'italic'))


plot_slope_gen
```

Combine

```{r}
plot_slope_both_gen <- plot_dhw + plot_slope_gen + plot_layout(widths = c(2, 1))

plot_slope_both_gen
```

```{r}
ggsave(
  filename = "plot_slope_both_gen.pdf",
  plot = plot_slope_both_gen,
  path = here("long_term/plots/"),
  width = 22,
  height = 16,
  units = "cm"
)
```

Get summary data (for info in text)

```{r}
slope_gen %>%
  group_by(genus) %>%
  summarise(
    median = median(slope_val * 100, na.rm = T),
    l_95 = quantile(slope_val * 100, 0.025, na.rm = T),
    u_95 = quantile(slope_val * 100, 0.975, na.rm = T)
  ) %>%
  mutate_if(is.numeric, txtRound, 2) %>%
  htmlTable(rnames = FALSE)
```
