---
title: "Clean transect data"
format: html
editor: visual
---

## Libraries

```{r}
library(tidyverse)
library(here)
library(brms)
library(readxl)
```

## Read data

Read data and clean column names

```{r}
dat <- read_excel(here("long_term/data/raw/20230926MPAMooreaData_substrate.xlsx"), 
                  sheet = "MPAMooreaData_substrate.txt", 
                  range = cell_cols("A:J")) %>% 
  janitor::clean_names()
```

## Check and clean data

Range of benthic cover

```{r}
dat$proportion %>% range()
```

Check names for `marine_area` s

```{r}
dat$marine_area %>% unique() %>% sort()
```

Check names `habitat`

```{r}
dat$habitat %>% unique() %>% sort()
```

Is `transect` a unique ID, i.e. it occurs only once per survey year and site?

```{r}
dat %>% 
  dplyr::select(-substrate, -proportion) %>% 
  distinct() %>% 
  group_by(year, marine_area, habitat, transect, season) %>% 
  summarise(n = n()) %>% 
  filter(n > 1)
```

Not always, check for each case why

**2009 Pihaena Barrier reef 1 Oct**

```{r}
dat %>% 
  mutate(row_number = row_number()) %>% 
  filter(year == 2009, 
         marine_area == "Pihaena", 
         habitat == "Barrier reef", 
         transect == 1, 
         season == "Oct")
```

Seems to be a typo. 2009-08-08 should be 2009-09-08 (consecutive row_numbers). Change manually

```{r}
dat$date[which(dat$date == as.Date("2009-08-08"))] <- as.Date("2009-09-08")#occurs only 1x
```

**2012 Tiahura Outer slope 1 Mar**

```{r}
dat %>% 
  mutate(row_number = row_number()) %>% 
  filter(year == 2012, marine_area == "Tiahura", habitat == "Outer slope", transect == 1, season == "Mar")
```

Typo. 2012-03-09 should be 2012-03-08 (consecutive row_numbers)

```{r}
dat$date[which(
  dat$date == as.Date("2012-03-09") & 
    dat$habitat == "Outer slope" & 
    dat$marine_area == "Tiahura" &  
    dat$transect == 1)] <- as.Date("2012-03-08")
```

**2016 Nuarei Outer slope 2 Mar**

```{r}
dat %>% 
  mutate(row_number = row_number()) %>% 
  filter(year == 2016, marine_area == "Nuarei", habitat == "Outer slope", transect == 2, season == "Mar")
```

Typo. 2016-04-26 should be 2016-04-19 (consecutive row_numbers)

```{r}
dat$date[which(
  dat$date == as.Date("2016-04-26") & 
    dat$transect == 2 & 
    dat$marine_area == "Nuarei" & 
    dat$habitat == "Outer slope")] <- as.Date("2016-04-19")
```

**2021 Motu Ahi Fringing reef 2 Mar**

```{r}
dat %>% 
  mutate(row_number = row_number()) %>% 
  filter(year == 2021, marine_area == "Motu Ahi", habitat == "Fringing reef", transect == 2, season == "Mar")

```

Change date 18074 to 2021-03-23 not 2021-03-18

```{r}
dat$date[18074] <- as.Date("2021-03-23")
```

**Is the sum of all proportions always 100%?**

```{r}
dat %>% 
  group_by(year, season, marine_area, habitat, transect) %>% 
  summarise(sum_perc = round(sum(proportion), 2)) %>% 
  filter(sum_perc != 1) %>% 
  arrange(year, season, marine_area, habitat, transect)
```

No. Check for each case why

**2021 Tiahura Outer slope 1 or 3\***

```{r}
dat %>% 
  mutate(row_number = row_number()) %>% 
  filter(year == 2021, marine_area == "Tiahura", habitat == "Outer slope", transect %in% c(1, 3))

```

Row 17615 should be transect 3

```{r}
dat$transect[17615] <- 3
```

**2016 Mar Tiahura Outer slope 1 or 3**

```{r}
dat %>% 
  mutate(row_number = row_number()) %>% 
  filter(year == 2016, marine_area == "Tiahura", habitat == "Outer slope", transect %in% c(1, 3))
```

```{r}
dat[13210:13220,] %>% view()
```

Row 13216 should be transect 1?

```{r}
dat$transect[13216] <- 1
```

**2009 Oct Motu Ahi Fringing reef 3**

```{r}
dat %>% 
  mutate(row_number = row_number()) %>% 
  filter(year == 2009, marine_area == "Motu Ahi", habitat == "Fringing reef", transect == 3, season == "Oct")
```

=\> No obvious typo?

**2010 Mar Haapiti Barrier reef 2**

```{r}
dat %>% 
  mutate(row_number = row_number()) %>% 
  filter(year == 2010, marine_area == "Haapiti", habitat == "Barrier reef", transect == 2, season == "Mar")
```

**2014 Mar Gendron Fringing reef 2**

```{r}
dat %>% 
  mutate(row_number = row_number()) %>% 
  filter(year == 2014, marine_area == "Gendron", habitat == "Fringing reef", transect == 2, season == "Mar")
```

**2015 Mar Afareaitu Outer slope 3**

```{r}
dat %>% 
  mutate(row_number = row_number()) %>% 
  filter(year == 2015, marine_area == "Afareaitu", habitat == "Outer slope", transect == 3, season == "Mar")
```

**2020 Mar Tiahura Outer slope 1**

```{r}
dat %>% 
  mutate(row_number = row_number()) %>% 
  filter(year == 2020, marine_area == "Tiahura", habitat == "Outer slope", transect == 1, season == "Mar")
```

checked Excel file, row 17404 should be t1 not t3

```{r}
dat$transect[17404] <- 1
```

**2021 Mar Gendron Outer slope 2**

```{r}
dat %>% 
  mutate(row_number = row_number()) %>% 
  filter(year == 2021, marine_area == "Gendron", habitat == "Outer slope", transect == 2, season == "Mar")
```

**2021 Mar Gendron Outer slope 3**

```{r}
dat %>% 
  mutate(row_number = row_number()) %>% 
  filter(year == 2021, marine_area == "Gendron", habitat == "Outer slope", transect == 3, season == "Mar")
```

**2021 Mar Motu Ahi Barrier reef 3**

```{r}
dat %>% 
  mutate(row_number = row_number()) %>% 
  filter(year == 2021, marine_area == "Motu Ahi", habitat == "Barrier reef", transect == 3, season == "Mar")
```

Check again 100% per survey and quadrat?

```{r}
dat %>% 
  group_by(year, season, marine_area, habitat, transect) %>% 
  summarise(sum_perc = round(sum(proportion), 2)) %>% 
  filter(sum_perc != 1) %>% 
  arrange(year, season, marine_area, habitat, transect)
```

Almost. not possible to resolve rest. ok for now

**Check if all groups have consecutive row numbers**

```{r}
dat %>% 
  mutate(row_number = row_number()) %>% 
  group_by(year, season, marine_area, habitat, transect) %>% 
  summarise(diff_rn = diff(row_number)) %>% 
  filter(diff_rn != 1)
  
```

**Does year always fit to date?**

```{r}
dat %>% 
  mutate(year2 = format(date, format= "%Y") %>% as.numeric()) %>% 
  filter(year2 != year)
```

yes

**Check groups**

```{r}
dat$substrate %>% unique() %>% sort()
```

Inconstant names, as Other, other, Stegastes turf, Stegastes Turf.

**Clean up names**

```{r}
dat <- dat %>% 
  mutate(substrate = ifelse(substrate == "Stegastes Turf", "Stegastes_turf", substrate),
         substrate = ifelse(substrate == "other", "Other", substrate),
         substrate = ifelse(substrate == "Other (anemone)", "Other_anemone", substrate),
         substrate = ifelse(substrate == "Corail mou", "Softcoral", substrate),
         substrate = ifelse(substrate == "Corail non identifié", "Other_coral", substrate),#no space
         substrate = ifelse(substrate == "pocillopora" , "Pocillopora", substrate),
         substrate = ifelse(substrate == "Dead coral" , "Dead_coral", substrate))#no space
```

**Clean up genera**

-   Napopora to Porites

-   only genera, no species

```{r}
dat <- dat %>% 
  mutate(substrate = ifelse(substrate == "Napopora", "Porites", substrate)) %>% 
  separate(col = substrate, into = "substrate", sep = " ") %>% #separate and keep genus name
  mutate(substrate = tolower(substrate)) %>% #just to be sure
  mutate(substrate = str_to_title(substrate)) %>% 
  group_by(across(c(-proportion))) %>% 
  summarise(proportion = sum(proportion, na.rm = T)) %>% 
  ungroup()
```

Make wide and long again to get data for all genera. Otherwise, for some surveys, no corals remain \> the survey would be excluded when sub-setting to corals. Like that, it's 0%

```{r}
dat <- dat %>% 
  pivot_wider(names_from = substrate, values_from = proportion, values_fill = 0) %>% 
  pivot_longer(cols = 9:ncol(.),names_to = "substrate", values_to = "proportion")
```

Assign benthic groups

```{r}
corals <- c("Acanthastrea",
            "Acropora",
            "Astrea",
            "Astreopora",
            "Coscinaraea",
            "Cyphastrea",
            "Fungia",
            "Gardineroseris",
            "Goniastrea",
            "Herpolitha",
            "Leptastrea",
            "Leptoseris",
            "Lobophyllia",
            "Montipora",
            "Other_coral",
            "Pavona",
            "Pocillopora",
            "Porites",
            "Psammocora",
            "Synarea")

macro_algae <- c("Asparagopsis",
                 "Boodlea",
                 "Caulerpa",
                 "Dictyota",
                 "Halimeda",
                 "Macroalgae",
                 "Padina",
                 "Sargassum",
                 "Turbinaria", 
                 "Peyssonnelia")

substrate_ <- c("Mud",
                "Pavement",
                "Rubble",
                "Sand")

other <- c("Millepora",
           "Other",
           "Softcoral",
           "Other_anemone")

cyanos <- c("Cyanophyceae")

turf <- c("Stegastes_turf")

dead_coral <- "Dead_coral"
```

```{r}
dat <- dat %>%
  mutate(benthic_group = case_when(substrate %in% corals ~ "coral",
                                   substrate %in% macro_algae ~ "macro_alga",
                                   substrate %in% substrate_ ~ "substrate",
                                   substrate %in% other ~ "other",
                                   substrate %in% turf ~ "turf",
                                   substrate %in% cyanos ~ "cyanos",
                                   substrate %in% dead_coral ~ "dead_coral"))




dat %>% 
  filter(is.na(benthic_group))
```

Is `transect` a unique ID, i.e. it occurs only once per survey year and site?

```{r}
dat %>% 
  dplyr::select(-substrate, -benthic_group, -proportion) %>% 
  distinct() %>% 
  group_by(year, season, marine_area, habitat, transect) %>% 
  summarise(n = n()) %>% 
  filter(n > 1)
```

Can data be grouped by date? I.e. all surveys per area/habitat/year/season at same date?

```{r}
dat %>% 
  group_by(year, marine_area, habitat, season) %>% 
  summarize(n_dates = length(unique(date))) %>% 
  filter(n_dates != 1)
```

Nope, not for year == 2012, marine_area == "Tiahura", habitat == "Outer slope"

```{r}
dat %>% 
  filter(year == 2012, marine_area == "Tiahura", habitat == "Outer slope") 
```

Manually reset to everything to 2012-03-09 (from 2012-03-08 for some)

```{r}
dat$date[
    dat$year == 2012 &
    dat$marine_area == "Tiahura" &
    dat$habitat == "Outer slope" &
    dat$date ==  as.Date("2012-03-08")] <- as.Date("2012-03-09")
```

### Plot data

```{r}
plot_mm_overview <- dat %>% 
  group_by(year, marine_area, habitat, season) %>% 
  summarize(n_transects = length(unique(transect))) %>% 
  ungroup() %>% 
  complete(year,  marine_area, habitat, season) %>% 
  mutate(n_transects = ifelse(is.na(n_transects), 0, n_transects)) %>% 
  ggplot(aes(x = year, y = n_transects))+
  geom_col(aes(y = 3), fill = "darkred")+
  geom_col(fill = "darkgreen")+
  geom_rect(aes(xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf),
            fill = NA, inherit.aes = FALSE,col='grey',alpha=.6) +
  geom_hline(yintercept = c(0:3), col = "white")+
  facet_grid(marine_area ~ habitat +  season)+
  labs(y = "# transects", caption = "Missing quadrats in red")+
  theme_minimal();plot_mm_overview
```

Filter out Atiha (Inner Bommie) and season Oct

```{r}
dat <- dat %>% 
  filter(season != "Oct", marine_area != "Atiha") %>% 
  dplyr::select(-season)
```

Replot replication

```{r}
plot_mm_overview_sel <- dat %>% 
  group_by(year, marine_area, habitat) %>% 
  summarize(n_transects = length(unique(transect))) %>% 
  ungroup() %>% 
  complete(year,  marine_area, habitat) %>% 
  mutate(n_transects = ifelse(is.na(n_transects), 0, n_transects)) %>% 
  ggplot(aes(x = year, y = n_transects))+
  geom_col(aes(y = 3), fill = "darkred")+
  geom_col(fill = "darkgreen")+
  geom_rect(aes(xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf),
            fill = NA, inherit.aes = FALSE,col='grey',alpha=.6) +
  geom_hline(yintercept = c(0:3), col = "white")+
  facet_grid(marine_area ~ habitat)+
  labs(y = "# transects", caption = "Missing quadrats in red")+
  theme_minimal();plot_mm_overview_sel
```

Nurari 2024

```{r}
dat %>% 
  mutate(row_number = row_number()) %>% 
  filter(year == 2023, marine_area == "Nuarei", habitat == "Fringing reef")
```

No transect 3??

### 100% per survey and quadrat for remaining data?

```{r}
dat %>% 
  group_by(year, marine_area, habitat, transect) %>% 
  summarise(sum_perc = round(sum(proportion), 2)) %>% 
  filter(sum_perc != 1) %>% 
  arrange(year, marine_area, habitat, transect)
```

```{r}
dat %>% 
  saveRDS(here("long_term/data/clean/clean_MooreaMPA.rds"))
```
