---
title: "Get summary data per year"
format: html
editor: visual
---

## Libraries

```{r}
library(tidyverse)
library(here)
library(janitor)
```

get data

```{r}
dat <- read.csv(here("long_term/data/raw/2025_01_30_surveys2024.csv")) %>% 
  clean_names() %>% 
  filter(habitat == "Outer slope") %>% 
  select(year, marine_area, habitat, transect, substrate, proportion)
```

Some checks

```{r}
dat$proportion %>% range()

dat$marine_area %>% unique() %>% sort()

dat$habitat %>% unique() %>% sort()

dat$substrate %>% unique() %>% sort()
```

Is `transect` a unique ID, i.e. it occurs only once per survey year and site?

```{r}
dat %>% 
  dplyr::select(year, marine_area, habitat, transect) %>% 
  distinct() %>% 
  group_by(year, marine_area, habitat, transect) %>% 
  summarise(n = n()) %>% 
  filter(n > 1)
```

**Clean up genera**

-   Napopora to Porites

-   only genera, no species

```{r}
dat <- dat %>% 
  mutate(substrate = ifelse(substrate == "Napopora", "Porites", substrate)) %>% 
  separate(col = substrate, into = "substrate", sep = " ") %>% #separate and keep genus name
  mutate(substrate = tolower(substrate)) %>% #just to be sure
  mutate(substrate = str_to_title(substrate)) %>% 
  group_by(across(c(-proportion))) %>% 
  summarise(proportion = sum(proportion, na.rm = T)) %>% 
  ungroup()
```

Assign benthic groups

```{r}
corals <- c("Acanthastrea",
            "Acropora",
            "Astrea",
            "Astreopora",
            "Coscinaraea",
            "Cyphastrea",
            "Fungia",
            "Gardineroseris",
            "Goniastrea",
            "Herpolitha",
            "Leptastrea",
            "Leptoseris",
            "Lobophyllia",
            "Montipora",
            "Other_coral",
            "Pavona",
            "Pocillopora",
            "Porites",
            "Psammocora",
            "Synarea")

macro_algae <- c("Asparagopsis",
                 "Boodlea",
                 "Caulerpa",
                 "Dictyota",
                 "Halimeda",
                 "Macroalgae",
                 "Padina",
                 "Sargassum",
                 "Turbinaria", 
                 "Peyssonnelia")

substrate_ <- c("Mud",
                "Pavement",
                "Rubble",
                "Sand")

other <- c("Millepora",
           "Other",
           "Softcoral",
           "Other_anemone")

cyanos <- c("Cyanophyceae")

turf <- c("Stegastes_turf")

dead_coral <- c("Dead_coral", "Dead")
```

```{r}
dat <- dat %>%
  mutate(benthic_group = case_when(substrate %in% corals ~ "coral",
                                   substrate %in% macro_algae ~ "macro_alga",
                                   substrate %in% substrate_ ~ "substrate",
                                   substrate %in% other ~ "other",
                                   substrate %in% turf ~ "turf",
                                   substrate %in% cyanos ~ "cyanos",
                                   substrate %in% dead_coral ~ "dead_coral"))




dat %>% 
  filter(is.na(benthic_group))
```

Get % vals for corals

```{r}
dat <- dat %>% 
  filter(benthic_group == "coral") %>% 
  group_by(year, transect, marine_area) %>% 
  summarise(perc = sum(100 * proportion))
```

Make wide and long again to get data for all genera. Otherwise, for some surveys, no corals remain \> the survey would be excluded when sub-setting to corals. Like that, it's 0%

```{r}
dat <- dat %>%
  pivot_wider(names_from = transect, values_from = perc, values_fill = 0) %>% 
  pivot_longer(cols = 3:ncol(.),names_to = "transect", values_to = "perc")
```

Check no of transects per site

```{r}
dat %>% 
  group_by(year, marine_area) %>% 
  summarise(n = n()) %>% 
  filter(n != 3)
#ok

```

Means per site

```{r}
dat <- dat %>% 
  group_by(year, marine_area) %>% 
  summarise(perc = mean(perc)) 
```

add orientation

```{r}
dat_meta <- read.csv(here("long_term/data/clean/moorea_mpa_site_metadata.csv"))

dat <- dat %>% 
  left_join(dat_meta, by = "marine_area")

dat %>% 
  filter(is.na(coast))
```

Means per orientation

```{r}
datS <- dat %>% 
  group_by(year, coast) %>% 
  summarise(
    mean_perc = mean(perc),
    sd = sd(perc),
    n = n(),
    se = sd / sqrt(n)
  ) %>% 
  select(-sd, -n) %>% 
  mutate(across(where(is.numeric), round, 1))
```
