---
title: "DHW MUR Download, calculation, and plots"
format: html
editor: visual
bibliography: references.bib
---

```{r}
library(httr) #for downloads
library(here) #for project folder structure
library(tidyverse)
library(pbapply) #for progress bars
library(raster) #to load .nc files
library(lubridate) #fast date change function
library(sf) # mapping
library(ggspatial) # mapping
library(viridis) #plots
library(patchwork) #plots
library(ncdf4) # read files
library(patchwork)
library(zoo)
library(ggrepel)
library(giscoR) # for the countries dataset only
```

Get functions that help to download DHW and SST values

```{r}
source(here("dhw/scripts/Andi_DHW_functions.R"))
```

Define Moorea ROI

```{r}
moorea_lat = -17.536273
moorea_lon = -149.832730

moorea_box = bounding_box(moorea_lat, moorea_lon, 15, 15)
```

Base map Moorea

```{r}
map_FrenchPolynesia <- st_read(dsn = here("dhw/data/maps/polynesia_map/PYF_adm0.shp")) %>%
  st_as_sf()
```

Area to be downloaded. Add MMM as overlay

```{r}
map_moorea <- ggplot() +
  geom_sf(data = map_FrenchPolynesia,
          fill = "antiquewhite",
          col = "black")+
  coord_sf(xlim = c(moorea_box$lon_min, moorea_box$lon_max),
           ylim = c(moorea_box$lat_min, moorea_box$lat_max))+
  labs(x=NULL, y = NULL, title = "Moorea")+
  scale_fill_viridis(name = "MMM")+
  theme_minimal()

map_moorea
```

Downlaod MUR SST Data

```{r}

#download more to calculate DHW

# download_all(lat1 = moorea_box$lat_min,
#              lat2 = moorea_box$lat_max,
#              lon1 = moorea_box$lon_min,
#              lon2 = moorea_box$lon_max,
#              day1 = as.Date("2019-01-01")-12*7,
#              day2 = "2019-12-31",
#              parameter = "MUR_SST",
#              folder = here("dhw/data/sst"))

```

```{r}
dat_moorea_mur_sst <- read_nc(here("dhw/data/sst"),
                      varname = "analysed_sst",
                      values_to = "mur_sst")
```

Calculate DHW based on literature MMM

-   DHW calculated with 29°C as bleaching threshold (from [@pratchett2013])

```{r}
dat_moorea_mur_sst_dhw <- dat_moorea_mur_sst %>% 
  group_by(lat, lon) %>% 
  mutate(hot_spot = mur_sst - 29) %>% 
  mutate(hot_spot = ifelse(hot_spot >= 0, hot_spot, 0)) %>% 
  group_by(lat,lon) %>% 
  arrange(date) %>%
  mutate(dhw = c(rep(NA, 12*7-1),
                 rollapply(data = hot_spot/7, width = 84, FUN = sum))) %>% 
  ungroup() 
```

Get max DHW

```{r}
dat_moorea_mur_sst_dhwS <- dat_moorea_mur_sst_dhw %>% 
  group_by(lat, lon) %>% 
  summarise(max_dhw = max(dhw, na.rm = T)) %>% 
  mutate_all(na_if, -Inf) %>% 
  ungroup()
```

Add max dhw to site coordinates

```{r}
dat_sites <- read.csv(here("dhw/data/dat_mm_site_coordinates.csv"))
```

```{r}
dat_sites <- dat_sites %>% 
  rowwise() %>% 
  mutate(max_dhw = find_closest_maxdhw(data = dat_moorea_mur_sst_dhwS, 
                                       lat = lat, 
                                       lon = lon, 
                                       "max_dhw")) %>% 
  ungroup()
```

```{r}
write.csv(dat_sites, file = here("dhw/data/dat_sites_maxdhw_mur.csv"),row.names = F)
```

Overview mean and se

```{r}
dat_sites %>% 
  group_by(coast) %>% 
  summarise(mean = mean(max_dhw),
            sd = sd(max_dhw),
            n = n(),
            se = round(sd / sqrt(n),2))

```

Map with sites and `max_dhw_29` values

Get map of coral reef distribution

```{r}
map_corals <- st_read(dsn = here("dhw/data/maps/Coralreef_distribution_polynesia/Coralreef_distribution_polynesia.shp"), quiet = T) %>% 
  st_as_sf()
```

Get site coordinates

```{r}
sites <- read.csv(here("dhw/data/dat_mm_site_coordinates.csv"))
```

Make map

```{r}
#define colours
col_coast = c("West" = "#0A9F9DFF", 
              "North" = "#CEB175FF", 
              "East" = "#E54E21FF")


#get site df and define positioning of label
sites <- sites %>% 
  mutate(hjust = ifelse(coast == "West", 1, 0),
         nudge_x = ifelse(coast == "West", -0.008, 0.005),
         nudge_x = ifelse(site %in% c("Tiahura", "Entre 2 baies"),-0.005 ,nudge_x),
         nudge_y = case_when(site == "Tema'e" ~ -0.002 ,
                             site == "Nuarei" ~ 0.004 ,
                             site == "Gendron" ~ 0.004 ,
                             site == "Tetaiuo" ~ -0.004 ,
                             site == "Piha'ena" ~ -0.004 ,
                             site == "Tiahura" ~ 0.004,
                             site == "Entre 2 baies" ~ 0.004,
                             TRUE ~ 0),
         hjust = ifelse(site %in% c("Tiahura", "Entre 2 baies"),1 ,hjust)) 



#define outlines
lat_min <- -17.62
lat_max <- -17.46
lon_min <- -149.94
lon_max <- -149.74


plot_dhw_map <- ggplot() + 
  geom_raster(data = dat_moorea_mur_sst_dhwS %>% 
                filter(lat >= lat_min,
                       lat <= lat_max,
                       lon >= lon_min,
                       lon <= lon_max), 
              aes(x = lon, y = lat, fill = max_dhw)) +
  geom_sf(data = map_corals, fill = "white", col = NA, alpha = .4)+ # outlines reefs
  geom_sf(data = map_FrenchPolynesia,
          fill = "antiquewhite", #fill colour of land
          col = "black")+ # outlines
  #outline without hassle of having 2 fill legends
    geom_point(data = sites, 
             aes(x = lon, y = lat), 
             size = 2.4, 
             col = "white",
            show.legend = F)+
  #actual coloured points
  geom_point(data = sites, 
             aes(x = lon, y = lat, col = coast), 
             size = 2, 
            show.legend = F)+
  geom_text_repel(data = sites, 
            aes(x = lon, 
                y = lat, 
                label = site,
                hjust = hjust,
                col = coast),
             nudge_x = sites$nudge_x,
             nudge_y = sites$nudge_y,
            size = 3.3, 
            fontface = "italic",
            show.legend = F,
            min.segment.length = 2,
            force_pull = 2, 
            seed = 12,
            point.size = NA,
            bg.color = "white")+
  geom_text(aes(x = -149.85, y = -17.538, label = "Mo'orea"), 
            size = 4, 
            hjust = 0 #put label right of point
            )+
  annotation_scale(location = "br") + # scale at bottom right
  scale_fill_viridis(name = expression(AHS[max]))+
  scale_y_continuous(breaks = seq(-17.7, -17.4, by = 0.1)) + 
  scale_x_continuous(breaks = seq(-150, -149.7, by = 0.1)) +
  coord_sf(xlim = c(lon_min, lon_max),
           ylim = c(lat_min, lat_max),expand = F)+
  scale_colour_manual(values = col_coast, name = NULL)+
  theme_minimal()+
  labs(x = NULL, y = NULL)+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        strip.background = element_blank(),
        axis.ticks = element_line(colour = "black" ))+
  theme(legend.position = "top")

plot_dhw_map





ggsave("plot_dhw_map.png", 
       plot = plot_dhw_map, 
       path = here("dhw/plots"), 
       width = 15, height = 14, units = "cm",
       dpi = 500)

```

Overview World map

```{r}
crs_string <- "+proj=ortho +lon_0=-150 +lat_0=-17"

ocean <- st_point(x = c(0,0)) %>%
  st_buffer(dist = 6371000) %>%
  st_sfc(crs = crs_string)

world <- gisco_countries %>%
  st_intersection(ocean %>% st_transform(4326)) %>% # select visible area only
  st_transform(crs = crs_string) # reproject to ortho




world_map <- ggplot() + 
  geom_sf(data = ocean, fill = "aliceblue", color = NA)+
  geom_sf(data = world, fill = "grey", col = "grey")+
  geom_point(aes(x= -149.7598, y = -17.50671), shape = 1, col = "darkred", size = 10)+
  theme_void()

world_map

ggsave("world_map.pdf", plot = world_map, path = here("dhw/plots"), width = 10, height = 10, units = "cm")
```
